---
title: "CaseStudy2"
author: "Chris Roche"
date: "2/18/2021"
output: html_document
---

```{r, echo=FALSE}
suppressMessages( library(tidyverse) )
suppressMessages( library(class) )   # needed for knn
suppressMessages( library(dplyr) )
suppressMessages( library(caret) )   # needed for confusionMatrix
suppressMessages( library(ggthemes) )
suppressMessages( library(GGally) )
suppressMessages( library(plotly) )
suppressMessages( library(Metrics))
suppressMessages( library(readxl) )
suppressMessages( library(ggpubr) )

```

# Evaluating Attrition and Predicting Monthly Income
<br />  
My name is Chris Roche. In this file you'll find my analysis of attrition and monthly income data from your company. Using the data set provided to my organization, I have conducted exploratory data analysis, implemented a K-Nearest Neighbor model for classifying an employee's attrition, and a linear regression model for prediction an employees monthly income based on their years in industry.
<br />  
Here's a link to my presentation of this report: https://youtu.be/2mG4sgvjgwM
<br />  
  
#### Load in files
First we will load in three files:  
* the training data  
* unlabeled test data for attrition classification  
* unlabeled test data for monthly salary prediction  
```{r}
caseStudy2data = read.csv("data/CaseStudy2-data.csv", header = TRUE)
caseStudy2noAtt = read.csv("data/CaseStudy2CompSet-NoAttrition.csv", header = TRUE)
caseStudy2noSalary = read_excel("data/CaseStudy2CompSet-NoSalary.xlsx")
```
<br />  

#### Clean data we'll use for training
Begin by turning specific variables into factors with discrete levels. Check for missing data.
```{r}
caseStudy2data$Attrition <- as.factor(caseStudy2data$Attrition)
caseStudy2data$BusinessTravel <- as.factor(caseStudy2data$BusinessTravel)
caseStudy2data$Department <- as.factor(caseStudy2data$Department)
caseStudy2data$EducationField <- as.factor(caseStudy2data$EducationField)
caseStudy2data$Gender <- as.factor(caseStudy2data$Gender)
caseStudy2data$MaritalStatus <- as.factor(caseStudy2data$MaritalStatus)
caseStudy2data$Over18 <- as.factor(caseStudy2data$Over18)
caseStudy2data$OverTime <- as.factor(caseStudy2data$OverTime)
caseStudy2data$JobRole <- as.factor(caseStudy2data$JobRole)

# Look for missing values -- there are none
row.has.na <- apply(caseStudy2data, 1, function(x){any(is.na(x))})
sum(row.has.na)
```
The zero above indicates that there are no missing values within the data set provided.
<br />  

## Part 1 - KNN for Attrition of emplyees
#### Begin exploring correlation of several variables for Attrition
Below are the 3 variables identifed for highest Sensitivity/Specificity  
* Employee's Job level (1 - 5; 5 being highest)  
* Number of companies worked for in employee's career  
* Number of years employed by your company  
```{r}
# 1 - Job Level
caseStudy2data %>% ggplot(aes(x = JobLevel, y = Attrition)) + geom_boxplot(fill="violetred4") + ggtitle("Job Level vs Attrition") + theme_minimal()

p = caseStudy2data %>% 
  count(JobLevel,Attrition) %>% 
  ggplot(aes(x=JobLevel,y=Attrition)) + geom_tile(mapping = aes(fill = n)) + ggtitle("Job Level vs Attrition") + theme_minimal()
ggplotly(p)

# 2 - Number of Companies worked
caseStudy2data %>% ggplot(aes(x = NumCompaniesWorked, y = Attrition)) + geom_boxplot(fill="violetred4") + ggtitle("Number of Companies Worked vs Attrition") + theme_minimal()

p = caseStudy2data %>% 
  count(NumCompaniesWorked,Attrition) %>% 
  ggplot(aes(x=NumCompaniesWorked,y=Attrition)) + geom_tile(mapping = aes(fill = n)) + ggtitle("Number of Companies Worked vs Attrition") + theme_minimal()
ggplotly(p)

# 3 - Years at the company
caseStudy2data %>% ggplot(aes(x = YearsAtCompany, y = Attrition)) + geom_boxplot(fill="violetred4") + ggtitle("Years at Company vs Attrition") + theme_minimal()

# Mean and Median Years at Company
caseStudy2data %>%
  group_by(Attrition) %>%
    summarise(MeanYearsAtCompany=mean(YearsAtCompany),
              MedianYearsAtCompany=median(YearsAtCompany))
```
<br />  

#### Prepare the data
Divide up the labeled employee data into a 70/30 train/test split.
```{r}
set.seed(4)
splitPerc = .7

# pull out 3 explanatory variables and attrition
caseStudy2SelectData = caseStudy2data %>% 
    select(JobLevel, NumCompaniesWorked, YearsAtCompany, Attrition) #87,67

# use split percentage above to split into train and test
trainIndices = sample(1:dim(caseStudy2SelectData)[1], round(splitPerc * dim(caseStudy2SelectData)[1]))

# put my train indices into a train df
caseStudyTrain = caseStudy2SelectData[trainIndices,]

# put my NOT (-) train indices into a test df
caseStudyTest = caseStudy2SelectData[-trainIndices,]
```
<br />  

#### Implement KNN Model for Attrition
This model uses the three contributors identified above to classify an employee's attrition status.    
  
Key results obtained:  
* Accuracy    : 87.4%  
* Sensitivity : 87.6%  
* Specificity : 66.7%  
```{r}
# run knn
classifications = knn(caseStudyTrain[,1:3], caseStudyTest[,1:3], caseStudyTrain$Attrition, k = 5, prob = TRUE)

# Confusion Matrix
table(caseStudyTest$Attrition, classifications)
confusionMatrix(table(caseStudyTest$Attrition, classifications))
```
<br />  

#### Run KNN on the provided test data and write predictions to file
Using the unlabeled data set read in above and the KNN model, classify the test observations and write back out to a file for post-processing.
File name: data/Case2PredictionsRoche_Attrition.csv
```{r}
caseStudy2SelectNoAtt = caseStudy2noAtt %>% 
    select(JobLevel, NumCompaniesWorked, YearsAtCompany, ID)

classifications = knn(caseStudyTrain[,1:3], caseStudy2SelectNoAtt[,1:3], caseStudyTrain$Attrition, k = 5, prob = TRUE)

# Cleanup the df to match example file before writing
dfKNN = as.data.frame(classifications)
dfKNN$ID <- caseStudy2SelectNoAtt$ID
dfKNN <- dfKNN %>% rename(Attrition = classifications)
dfKNN <- dfKNN[,c(2,1)]

# Write to file
write.csv(dfKNN, "data/Case2PredictionsRoche_Attrition.csv", row.names = FALSE, quote=FALSE)
```
<br />  

## Part 2 - Linear Regression for Monthly Income
#### Prepare the data
Divide up the labeled employee data into a 70/30 train/test split.
```{r}
# RMSE = sqrt of the sum of Predicted minus Actual squared, over N
set.seed(4)
splitPerc = .7

# Data set to use for LM
caseStudy2SelectLMData = caseStudy2data
caseStudy2SelectLMData = caseStudy2SelectLMData %>% mutate(TotalWorkingYears2 = TotalWorkingYears^2)
caseStudy2SelectLMData = caseStudy2SelectLMData %>% mutate(TotalWorkingYears3 = TotalWorkingYears^3)

# use split percentage above to split into train and test
trainIndicesLM = sample(1:dim(caseStudy2SelectLMData)[1], round(splitPerc * dim(caseStudy2SelectLMData)[1]))

# put my train indices into a train df
caseStudyLMTrain = caseStudy2SelectLMData[trainIndices,]

# put my NOT (-) train indices into a test df
caseStudyLMTest = caseStudy2SelectLMData[-trainIndices,]
```
<br />  

#### Run the model
Run a linear regression model using the following parameters:  
1. Total Working Years  
2. Total Working Years Squared  
3. Total Working Years Cubed  
```{r}
# fit the lm for the train
monthlyIncomeFit = lm(MonthlyIncome~TotalWorkingYears+TotalWorkingYears2+TotalWorkingYears3, data = caseStudyLMTrain)
summary(monthlyIncomeFit)
```
<br />  

##### Get predictions with test data
Using the model, run it against the test data
```{r, results='hide'}
monthlyIncomeFitPredictions = predict(monthlyIncomeFit, newdata = caseStudyLMTest)
as.data.frame(monthlyIncomeFitPredictions)
```
<br />  

##### Calculate RMSE using the predictions and truth data
Determine the Root Mean Squared Error of the predictions  
Performance achieved: < $2716  
```{r}
RMSE = data.frame(Actual = caseStudyLMTest$MonthlyIncome, Predicted = monthlyIncomeFitPredictions)
RMSE$Resisdual = RMSE$Predicted - RMSE$Actual
RMSE$SquaredResidual = RMSE$Resisdual^2
sqrt(mean(RMSE$SquaredResidual))
```
<br />  

#### Plot the regression line of the predicted data against the observations
```{r}
caseStudyLMTest %>% ggplot(aes(x = TotalWorkingYears, y = MonthlyIncome)) + 
  geom_point() + 
  geom_line(data = caseStudyLMTest, aes( x = TotalWorkingYears, 
      y = monthlyIncomeFitPredictions, col = "red")) + 
  stat_cor(label.y = 150, aes(label = paste(..rr.label.., 
      ..p.label.., sep = "~`,`~")))
```
<br />  

#### Run the model against the provided test data and save to a csv file
Using the unlabeled data set read in above and the linear regression model, classify the test observations and write back out to a file for post-processing.
File name: data/Case2PredictionsRoche_Salary.csv
```{r}
caseStudy2noSalary = caseStudy2noSalary %>% mutate(TotalWorkingYears2 = TotalWorkingYears^2)
caseStudy2noSalary = caseStudy2noSalary %>% mutate(TotalWorkingYears3 = TotalWorkingYears^3)
monthlyIncomeFitPredictions = predict(monthlyIncomeFit, newdata = caseStudy2noSalary)

# Cleanup the df to match example file before writing
dfLM = as.data.frame(monthlyIncomeFitPredictions)
dfLM$ID <- caseStudy2noSalary$ID
dfLM <- dfLM[,c(2,1)]
dfLM <- dfLM %>% rename(MonthlyIncome = monthlyIncomeFitPredictions)
dfLM$MonthlyIncome <- round(dfLM$MonthlyIncome)

# Write to file
write.csv(dfLM, "data/Case2PredictionsRoche_Salary.csv", row.names = FALSE, quote=FALSE)
```
<br />  

## Part 3 - Insights by job role
#### Plot by JobRole as a factor
A detailed look into median satisfaction ratings and attrition percentages based on job role.
```{r}
# Summarize the number of onbservations by Job Tole
summary(caseStudy2data$JobRole)


# Does one job role have higher environment satisfaction?
caseStudy2data %>%
  group_by(JobRole) %>%
    summarize(Environment=mean(EnvironmentSatisfaction),
              Job=mean(JobSatisfaction),
              Relationship=mean(RelationshipSatisfaction),
              Average=((mean(EnvironmentSatisfaction) + 
                          mean(JobSatisfaction) + 
                          mean(RelationshipSatisfaction))/3),
              Income=mean(MonthlyIncome))


# Does one job role have higher attrition?
caseStudy2data %>%
  group_by(JobRole) %>%
    summarise(Attrition.No = sum(Attrition == "No", na.rm=T) / n(), 
              Attrition.Yes = sum(Attrition == "Yes", na.rm=T) / n() )

```
<br />  

```{r, echo=FALSE, eval = FALSE}
## Extra


# Maybe
caseStudy2data %>% ggplot(aes(x = JobSatisfaction, y = Attrition)) + geom_boxplot() + ggtitle("Job Satisfaction vs Attrition")

# No
caseStudy2data %>% ggplot(aes(x = PerformanceRating, y = Attrition)) + geom_boxplot() + ggtitle("Performance Rating vs Attrition")

# No
caseStudy2data %>% ggplot(aes(x = YearsSinceLastPromotion, y = Attrition)) + geom_boxplot() + ggtitle("Years Since Last Promotion vs Attrition")

# Maybe
caseStudy2data %>% ggplot(aes(x = DistanceFromHome, y = Attrition)) + geom_boxplot() + ggtitle("Commute vs Attrition")

# No
caseStudy2data %>% ggplot(aes(x = JobInvolvement, y = Attrition)) + geom_boxplot() + ggtitle("Job Involvement vs Attrition")

# No
caseStudy2data %>% ggplot(aes(x = StockOptionLevel, y = Attrition)) + geom_boxplot() + ggtitle("Stock Option Level vs Attrition")

# No
caseStudy2data %>% ggplot(aes(x = WorkLifeBalance, y = Attrition)) + geom_boxplot() + ggtitle("Work Life Balance vs Attrition")

# Maybe
caseStudy2data %>% ggplot(aes(x = YearsInCurrentRole, y = Attrition)) + geom_boxplot() + ggtitle("Years in Current Role vs Attrition")

# Maybe
caseStudy2data %>% ggplot(aes(x = YearsWithCurrManager, y = Attrition)) + geom_boxplot() + ggtitle("Years with Current Manager vs Attrition")

# Maybe
caseStudy2data %>% ggplot(aes(x = Education, y = Attrition)) + geom_boxplot() + ggtitle("Education vs Attrition")

# Maybe
caseStudy2data %>% ggplot(aes(x = EnvironmentSatisfaction, y = Attrition)) + geom_boxplot() + ggtitle("Environment Satisfaction vs Attrition")

# No
caseStudy2data %>% ggplot(aes(x = PercentSalaryHike, y = Attrition)) + geom_boxplot() + ggtitle("Percent Salary Hike vs Attrition")

# No
caseStudy2data %>% ggplot(aes(x = TrainingTimesLastYear, y = Attrition)) + geom_boxplot() + ggtitle("Training vs Attrition")
```
<br />  
<br />  
Thank you! And don't forget to view my presentation: https://youtu.be/2mG4sgvjgwM  
<br />  