---
title: "CaseStudy2"
author: "Chris Roche"
date: "2/18/2021"
output: html_document
---

```{r, echo=FALSE}
suppressMessages( library(tidyverse) )
suppressMessages( library(class) )   # needed for knn
suppressMessages( library(dplyr) )
suppressMessages( library(caret) )   # needed for confusionMatrix
suppressMessages( library(ggthemes) )
suppressMessages( library(GGally) )
suppressMessages( library(plotly) )
suppressMessages( library(Metrics))
suppressMessages( library(readxl) )

```

# Evaluating Attrition and Predicting Monthly Income
<br />  

#### Load in files
```{r}
caseStudy2data = read.csv("data/CaseStudy2-data.csv", header = TRUE)
caseStudy2noAtt = read.csv("data/CaseStudy2CompSet-NoAttrition.csv", header = TRUE)
caseStudy2noSalary = read_excel("data/CaseStudy2CompSet-NoSalary.xlsx")
```
<br />  

#### Clean data we'll use for training
```{r}
caseStudy2data$Attrition <- as.factor(caseStudy2data$Attrition)
caseStudy2data$BusinessTravel <- as.factor(caseStudy2data$BusinessTravel)
caseStudy2data$Department <- as.factor(caseStudy2data$Department)
caseStudy2data$EducationField <- as.factor(caseStudy2data$EducationField)
caseStudy2data$Gender <- as.factor(caseStudy2data$Gender)
caseStudy2data$MaritalStatus <- as.factor(caseStudy2data$MaritalStatus)
caseStudy2data$Over18 <- as.factor(caseStudy2data$Over18)
caseStudy2data$OverTime <- as.factor(caseStudy2data$OverTime)

# Look for missing values -- there are none
row.has.na <- apply(caseStudy2data, 1, function(x){any(is.na(x))})
sum(row.has.na)
```
<br />  

## Part 1 - KNN for Attrition of emplyees
#### Begin exploring correlation of several variables for Attrition
#### Below are the 3 variables identifed for highest Sensitivity/Specificity
```{r}
# 1 - Job Level
caseStudy2data %>% ggplot(aes(x = JobLevel, y = Attrition)) + geom_boxplot() + ggtitle("Job Level vs Attrition")

p = caseStudy2data %>% 
  count(JobLevel,Attrition) %>% 
  ggplot(aes(x=JobLevel,y=Attrition)) + geom_tile(mapping = aes(fill = n)) + ggtitle("Job Level vs Attrition")
ggplotly(p)

# 2 - Number of Companies worked
caseStudy2data %>% ggplot(aes(x = NumCompaniesWorked, y = Attrition)) + geom_boxplot() + ggtitle("Number of Companies Worked vs Attrition")

p = caseStudy2data %>% 
  count(NumCompaniesWorked,Attrition) %>% 
  ggplot(aes(x=NumCompaniesWorked,y=Attrition)) + geom_tile(mapping = aes(fill = n)) + ggtitle("Number of Companies Worked vs Attrition")
ggplotly(p)

# 3 - Years at the company
caseStudy2data %>% ggplot(aes(x = YearsAtCompany, y = Attrition)) + geom_boxplot() + ggtitle("Years at Company vs Attrition")

```
<br />  

#### Implement KNN Model for Attrition
```{r}
set.seed(4)
splitPerc = .7

# pull out 3 explanatory variables and attrition
caseStudy2SelectData = caseStudy2data %>% 
    select(JobLevel, NumCompaniesWorked, YearsAtCompany, Attrition) #87,67
    #select(NumCompaniesWorked, YearsAtCompany, YearsInCurrentRole, Attrition) #87,50
    #select(YearsInCurrentRole, YearsWithCurrManager, Education, Attrition) #87,50

# use split percentage above to split into train and test
trainIndices = sample(1:dim(caseStudy2SelectData)[1], round(splitPerc * dim(caseStudy2SelectData)[1]))

# put my train indices into a train df
caseStudyTrain = caseStudy2SelectData[trainIndices,]

# put my NOT (-) train indices into a test df
caseStudyTest = caseStudy2SelectData[-trainIndices,]

# run knn
classifications = knn(caseStudyTrain[,1:3], caseStudyTest[,1:3], caseStudyTrain$Attrition, k = 5, prob = TRUE)

# Confusion Matrix
table(caseStudyTest$Attrition, classifications)
confusionMatrix(table(caseStudyTest$Attrition, classifications))
```
<br />  

#### Run KNN on the provided test data and write predictions to file
```{r}
caseStudy2SelectNoAtt = caseStudy2noAtt %>% 
    select(JobLevel, NumCompaniesWorked, YearsAtCompany, ID)

classifications = knn(caseStudyTrain[,1:3], caseStudy2SelectNoAtt[,1:3], caseStudyTrain$Attrition, k = 5, prob = TRUE)

# Cleanup the df to match example file before writing
dfKNN = as.data.frame(classifications)
dfKNN$ID <- caseStudy2SelectNoAtt$ID
dfKNN <- dfKNN %>% rename(Attrition = classifications)
dfKNN <- dfKNN[,c(2,1)]

# Write to file
write.csv(dfKNN, "data/Case2PredictionsRoche_Attrition.csv", row.names = FALSE, quote=FALSE)
```
<br />  

## Part 2 - Linear Regression for Monthly Income
#### Prepare the data
```{r}
# RMSE = sqrt of the sum of Predicted minus Actual squared, over N
set.seed(4)
splitPerc = .7

# Data set to use for LM
caseStudy2SelectLMData = caseStudy2data
caseStudy2SelectLMData = caseStudy2SelectLMData %>% mutate(TotalWorkingYears2 = TotalWorkingYears^2)
caseStudy2SelectLMData = caseStudy2SelectLMData %>% mutate(TotalWorkingYears3 = TotalWorkingYears^3)

# use split percentage above to split into train and test
trainIndicesLM = sample(1:dim(caseStudy2SelectLMData)[1], round(splitPerc * dim(caseStudy2SelectLMData)[1]))

# put my train indices into a train df
caseStudyLMTrain = caseStudy2SelectLMData[trainIndices,]

# put my NOT (-) train indices into a test df
caseStudyLMTest = caseStudy2SelectLMData[-trainIndices,]
```
<br />  

#### Run the model and calculate RMSE
```{r}
# fit the lm for the train
monthlyIncomeFit = lm(MonthlyIncome~TotalWorkingYears+TotalWorkingYears2+TotalWorkingYears3, data = caseStudyLMTrain)
summary(monthlyIncomeFit)
```
<br />  

##### Get predictions with test data
```{r, results='hide'}
monthlyIncomeFitPredictions = predict(monthlyIncomeFit, newdata = caseStudyLMTest)
as.data.frame(monthlyIncomeFitPredictions)
```
<br />  

##### Manually calculate RMSE using the predictions and truth data
```{r}
RMSE = data.frame(Actual = caseStudyLMTest$MonthlyIncome, Predicted = monthlyIncomeFitPredictions)
RMSE$Resisdual = RMSE$Predicted - RMSE$Actual
RMSE$SquaredResidual = RMSE$Resisdual^2
sqrt(mean(RMSE$SquaredResidual))
```
<br />  

#### Plot the regression line of the predicted data against the observations
```{r}
caseStudyLMTest %>% ggplot(aes(x = TotalWorkingYears, y = MonthlyIncome)) + geom_point() + geom_line(data = caseStudyLMTest, aes( x = TotalWorkingYears, y = monthlyIncomeFitPredictions, col = "red"))

```
<br />  

#### Run model against the provided test data and save to a csv file
```{r}
caseStudy2noSalary = caseStudy2noSalary %>% mutate(TotalWorkingYears2 = TotalWorkingYears^2)
caseStudy2noSalary = caseStudy2noSalary %>% mutate(TotalWorkingYears3 = TotalWorkingYears^3)
monthlyIncomeFitPredictions = predict(monthlyIncomeFit, newdata = caseStudy2noSalary)

# Cleanup the df to match example file before writing
dfLM = as.data.frame(monthlyIncomeFitPredictions)
dfLM$ID <- caseStudy2noSalary$ID
dfLM <- dfLM[,c(2,1)]
dfLM <- dfLM %>% rename(MonthlyIncome = monthlyIncomeFitPredictions)
dfLM$MonthlyIncome <- round(dfLM$MonthlyIncome)

# Write to file
write.csv(dfLM, "data/Case2PredictionsRoche_Salary.csv", row.names = FALSE, quote=FALSE)
```


```{r}

```


```{r, echo=FALSE, eval = FALSE}
## Extra


# Maybe
caseStudy2data %>% ggplot(aes(x = JobSatisfaction, y = Attrition)) + geom_boxplot() + ggtitle("Job Satisfaction vs Attrition")

# No
caseStudy2data %>% ggplot(aes(x = PerformanceRating, y = Attrition)) + geom_boxplot() + ggtitle("Performance Rating vs Attrition")

# No
caseStudy2data %>% ggplot(aes(x = YearsSinceLastPromotion, y = Attrition)) + geom_boxplot() + ggtitle("Years Since Last Promotion vs Attrition")

# Maybe
caseStudy2data %>% ggplot(aes(x = DistanceFromHome, y = Attrition)) + geom_boxplot() + ggtitle("Commute vs Attrition")

# No
caseStudy2data %>% ggplot(aes(x = JobInvolvement, y = Attrition)) + geom_boxplot() + ggtitle("Job Involvement vs Attrition")

# No
caseStudy2data %>% ggplot(aes(x = StockOptionLevel, y = Attrition)) + geom_boxplot() + ggtitle("Stock Option Level vs Attrition")

# No
caseStudy2data %>% ggplot(aes(x = WorkLifeBalance, y = Attrition)) + geom_boxplot() + ggtitle("Work Life Balance vs Attrition")

# Maybe
caseStudy2data %>% ggplot(aes(x = YearsInCurrentRole, y = Attrition)) + geom_boxplot() + ggtitle("Years in Current Role vs Attrition")

# Maybe
caseStudy2data %>% ggplot(aes(x = YearsWithCurrManager, y = Attrition)) + geom_boxplot() + ggtitle("Years with Current Manager vs Attrition")

# Maybe
caseStudy2data %>% ggplot(aes(x = Education, y = Attrition)) + geom_boxplot() + ggtitle("Education vs Attrition")

# Maybe
caseStudy2data %>% ggplot(aes(x = EnvironmentSatisfaction, y = Attrition)) + geom_boxplot() + ggtitle("Environment Satisfaction vs Attrition")

# No
caseStudy2data %>% ggplot(aes(x = PercentSalaryHike, y = Attrition)) + geom_boxplot() + ggtitle("Percent Salary Hike vs Attrition")

# No
caseStudy2data %>% ggplot(aes(x = TrainingTimesLastYear, y = Attrition)) + geom_boxplot() + ggtitle("Training vs Attrition")
```



<br />  
<br />  

End